---
title: "Render"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Render}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(docorator)
```

Running `as_docorator()` does not render your display - it creates a `docorator` object to store all your display information. For rendering, an explicit call to a render function is needed. A display can be output to PDF, RTF, or both via a combination of the `render_pdf()` and `render_rtf()` functions.

Further, with `render_pdf()`, users have a choice of R Markdown or Quarto engines via the `quarto` option. Currently, the result is nearly identical, except for {gt} tables, where `quarto = FALSE` renders the `gt` via the `longtable` LaTeX library and `quarto = TRUE` via the `tabular` LaTeX library. Please note the Quarto engine is currently experimental and can take longer to render. 

## Render to PDF

```
as_docorator(
  mytbl,
  display_name = "mytbl"
) |> 
render_pdf()
```

If you understand the underlying LaTeX involved in going from a display object to PDF, `render_pdf` has a few arguments which could be of use. 

- The `transform` argument allows you to pass a function to manipulate the LaTeX string of the object before it goes to PDF. 
- The `header_latex` argument allows additional LaTeX to be added to the header-includes section of the template.Rmd used for PDF generation. For example, you can pass additional LaTeX libraries here. This should be passed as a .tex file. 
- The `escape_latex` argument controls escaping for special characters in headers and footers. This can be set to `FALSE` to disable the escaping if manual escaping is preferred, i.e. if results are not as expected.

## Render to RTF

```
as_docorator(
  mytbl,
  display_name = "mytbl"
) |> 
render_rtf()
```

Due to the limitations of the RTF file format, some arguments specified in your `as_docorator()` call are not preserved, such as certain formatting or table styling. Document decoration headers are also not preserved, only titles specified as 'center' will be applied. 

## Render to HTML

```
as_docorator(
  mytbl,
  display_name = "mytbl"
) |> 
render_html()
```

`render_html()` writes a self-contained `.html` file, rendering the display object together with any headers and footers using HTML and CSS in a page-like display. Long tables are presented in a scrollable container with sticky headers in lieu of pagination. This render engine is experimental and the layout is likely to evolve over time. 

## Render to PDF via HTML
```
as_docorator(
  mytbl,
  display_name = "mytbl"
) |> 
render_pdf_html()
```

`render_pdf_html()` produces a PDF by first rendering the display to HTML (via `render_html()`) and then converting that HTML to PDF using [chromote](https://rstudio.github.io/chromote/). This conversion preserves the original styling more closely than the LaTeX-based `render_pdf()` pathway.

Please note `render_pdf_html()` depends on a Chrome/Chromium installation being available to chromote.

## Multi-render

```
as_docorator(
  mytbl,
  display_name = "mytbl"
) |>
render_pdf() |>
render_rtf()
```

Both `render_pdf()` and `render_rtf()` return the original object created with `as_docorator()`, and as such, can be piped one after another as shown in the example above.
